cubble An R Package for Structuring Spatio-temporal Data

### 2022-04-23 2-0-cube.Rmd
spell check

Spatio-temporal data can be conceptualised using a cubical data model with three axis. This abstraction can be useful to generalise operations or even visualisation: @lu_multidimensional_2018 shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations. Notice that the term space-time cube in their article "does not need to involve spatial data", but referring to "an abstract 2D substrate that is used to visualize data at a specific time". Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.

While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array,  may not be convenient for data wrangling. There are two reasons for this: 1) While arrays are efficient for computation on numerical values, spatio-temporal data contains more than just numerical variables. Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays. 2) The `mutate()` syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.

There has been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stack on time. Our spatio-temporal cube is defined slightly different with three axes: time, site, and variable  as Figure \ref{fig:cube-def}. The time axis is the same across the two formation, site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas. This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.

>>>>

Spatio-temporal data can be conceptualised using a cubical data model with three axes. This abstraction can be useful to generalise operations or even visualisation: @lu_multidimensional_2018 shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; @bach_review_2014 reviews the temporal data visualisation based on space-time cube operations. Notice that the term space-time cube in their article "does not need to involve spatial data", but refers to "an abstract 2D substrate that is used to visualize data at a specific time". Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.

While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array,  may not be convenient for data wrangling. There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables. Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays. 2) The `mutate()` syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.

There have been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time. Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}. The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas. This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.

### 2022-04-23 2-the-cubble-pkg.Rmd
spell check

Each core function is introduced with a short example accompanied and the data `climate_flat` is used throughout this section. `climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020. Below prints out the first five rows of `climate_flat`:

@@

The created cubble is built from a \code{rowwise_df} class where each row forms a group. All the temporal variables are nested in a list column, hence it is also called the nested cubble. The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simplier than the {\code{map()} when working with list column. For example, calculating the number of raining day can be done by:

@@

The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate. \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hide" the spatial variables. The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrate this operation where the focus of the cube now changes from the site-variable face to the time-variable face. This code switches the cubble just created into its long form:

@@

The first line in the header now shows the cubble is in its long form and the third line has been changed to show the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}. Unlike the nested form buildin form a \code{rowwise_df}, the long cubble is built form class \code{groupped_df} where all the observations from the same sites forms a group.

@@

Wrangling spatio-temproal data can be an iterative process on the spatial and temporal dimension. Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}. The second row of Figure \ref{fig:face} illustrates  the function and below is an example on the climate data:

@@

Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable. An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}. This type of operations can be seen as flattening, or *unfolding*, the cube into a 2D data frame. Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:

@@

Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column. While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data. While `nest_by()` can recreate the nested form, it can't create a clean slate for temporal variables as the long form. For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with. This is because these spatial variables will lost after \code{summarise()} and one will need to subset the spatial variables again from the origin data and make a left join, which tends to create unnecessary intermediate products in the workflow.

@@

As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc). An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble. Once formed in a cubble, methods for the \code{sfc} class can be applied on the nested form.  An example of this is shown in \@ref(covid), which also handles the case when the site in the two tables don't match exactly. A spatial data object with \code{s2} vector can also be the input of the spatial table.

@@

NetCDF data is another format that is commonly used for storing spatio-temporal data. It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes. A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

@@

The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution. The spatial grid can also be subsetted for to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:

>>>>

Each core function is introduced with a short example accompanied and the data `climate_flat` is used throughout this section. `climate_flat` contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020. Below prints out the first five rows of `climate_flat`:

@@

The created cubble is built from a \code{rowwise_df} class where each row forms a group. All the temporal variables are nested in a list column, hence it is also called the nested cubble. The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the {\code{map()} when working with list column. For example, calculating the number of raining days can be done by:

@@

The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate. \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily "hide" the spatial variables. The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face. This code switches the cubble just created into its long form:

@@

The first line in the header now shows the cubble is in its long form and the third line has been changed to show the name and type of spatial variables: \code{lat [dbl], long [dbl], elev [dbl], name [chr], wmo_id [dbl]}. Unlike the nested form, the long cubble is built form class \code{groupped_df} where all the observations from the same sites form a group.

@@

Wrangling spatio-temporal data can be an iterative process in the spatial and temporal dimensions. Switching the focus back to the site-variable face can be accomplished by the function \code{face_spatial()}, which is the inverse of \code{face_temporal()}. The second row of Figure \ref{fig:face} illustrates  the function and below is an example on the climate data:

@@

Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables. An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}. This type of operation can be seen as flattening, or *unfolding*, the cube into a 2D data frame. Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:

@@

Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column. While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data. While `nest_by()` can recreate the nested form, it can't create a clean slate for temporal variables as the long form. For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with. This is because these spatial variables will lose after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.

@@

As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc). An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble. Once formed in a cubble, methods for the \code{sfc} class can be applied in the nested form.  An example of this is shown in \@ref(covid), which also handles the case when the sites in the two tables don't match exactly. A spatial data object with \code{s2} vector can also be the input of the spatial table.

@@

NetCDF data is another format that is commonly used for storing spatio-temporal data. It has two main components: *dimension* for defining the spatio-temporal grid (longitude, latitude, and time) and *variable* that populates the defined grid. Attributes are usually associated with dimensions and variables in the NetCDF format data and a [metadata convention for climate and forecast](http://cfconventions.org/) has been designed to standardise the format of the attributes. A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} [@ncdf4], a low-level interface that calls C-interface: \pkg{RNetCDF} [@rnetcdf; @michna2013rnetcdf], and a tidyverse implementation: \pkg{tidync} [@tidync].

@@

The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year. A 300 by 300 spatial grid can be a bounding box of [100, -80, 180, 0] at 0.25 degree resolution or a global bounding box [-180, -90, 180, -90] at 1 degree resolution. The spatial grid can also be subsetted to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:

### 2022-04-23 arxiv/cubble-arxiv.tex
spell check

Spatio-temporal data can be conceptualised using a cubical data model with three axis. This abstraction can be useful to generalise operations or even visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al. (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations. Notice that the term space-time cube in their article ``does not need to involve spatial data'', but referring to ``an abstract 2D substrate that is used to visualize data at a specific time''. Despite its main focus is on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.

While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array, may not be convenient for data wrangling. There are two reasons for this: 1) While arrays are efficient for computation on numerical values, spatio-temporal data contains more than just numerical variables. Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays. 2) The \texttt{mutate()} syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.

There has been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stack on time. Our spatio-temporal cube is defined slightly different with three axes: time, site, and variable as Figure \ref{fig:cube-def}. The time axis is the same across the two formation, site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas. This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.

@@

Each core function is introduced with a short example accompanied and the data \texttt{climate\_flat} is used throughout this section. \texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020. Below prints out the first five rows of \texttt{climate\_flat}:

@@

The created cubble is built from a \code{rowwise_df} class where each row forms a group. All the temporal variables are nested in a list column, hence it is also called the nested cubble. The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simplier than the \{\code{map()} when working with list column. For example, calculating the number of raining day can be done by:

@@

The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate. \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hide'' the spatial variables. The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrate this operation where the focus of the cube now changes from the site-variable face to the time-variable face. This code switches the cubble just created into its long form:

@@

Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variable. An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}. This type of operations can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame. Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:

@@

Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column. While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data. While \texttt{nest\_by()} can recreate the nested form, it can't create a clean slate for temporal variables as the long form. For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with. This is because these spatial variables will lost after \code{summarise()} and one will need to subset the spatial variables again from the origin data and make a left join, which tends to create unnecessary intermediate products in the workflow.

@@

As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc). An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble. Once formed in a cubble, methods for the \code{sfc} class can be applied on the nested form. An example of this is shown in \ref{covid}, which also handles the case when the site in the two tables don't match exactly. A spatial data object with \code{s2} vector can also be the input of the spatial table.

@@

NetCDF data is another format that is commonly used for storing spatio-temporal data. It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid. Attributes are usually associated with dimension and variable in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes. A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).

@@

The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year. A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution. The spatial grid can also be subsetted for to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:

@@

Spatial locations can have grouping structure either inherent to the data or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas. In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial locations by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form. By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}. All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.

>>>>

Spatio-temporal data can be conceptualised using a cubical data model with three axes. This abstraction can be useful to generalise operations or even visualisation: Lu, Appel, and Pebesma (\protect\hyperlink{ref-lu_multidimensional_2018}{2018}) shows an example of how array operations (select, scale, reduce, rearrange, and compute) can be mapped onto the cube; Bach et al. (\protect\hyperlink{ref-bach_review_2014}{2014}) reviews the temporal data visualisation based on space-time cube operations. Notice that the term space-time cube in their article ``does not need to involve spatial data'', but refers to ``an abstract 2D substrate that is used to visualize data at a specific time''. Despite its main focus being on temporal data, the mindset of abstracting out data representation to construct visualisation is applicable to our spatio-temporal data manipulation and visualisation.

While a cubical conceptual data model is neat to conceptually view the spatio-temporal data, a 3D data representation, or an array, may not be convenient for data wrangling. There are two reasons for this: 1) While arrays are efficient for the computation on numerical values, spatio-temporal data contains more than just numerical variables. Wrangling on character strings and specific datetime classes are common tasks for spatio-temporal data but are not easy with arrays. 2) The \texttt{mutate()} syntax for creating new variables is no longer valid since the operation can no longer define how a created 1D vector should be added to the array.

There have been many cubes defined using time, latitude, and longitude, which can be seen as taking snapshots of the space and stacking on time. Our spatio-temporal cube is defined slightly differently with three axes: time, site, and variable as Figure \ref{fig:cube-def}. The time axis is the same across the two formations, the site encompasses latitude and longitude into a single column identifier, while variables are stacked on this defined space-time canvas. This notion is adopted to avoid the use of hypercube when describing multivariate spatio-temporal data.

@@

Each core function is introduced with a short example accompanied and the data \texttt{climate\_flat} is used throughout this section. \texttt{climate\_flat} contains five weather stations in Australia with spatial information of each station: station id, latitude, longitude, elevation, station name, and World Meteorology Organisation ID and also daily temporal information of date, maximum and minimum temperature and precipitation for 2020. Below prints out the first five rows of \texttt{climate\_flat}:

@@

The created cubble is built from a \code{rowwise_df} class where each row forms a group. All the temporal variables are nested in a list column, hence it is also called the nested cubble. The rowwise structure makes it simpler to calculate on the list using the \code{mutate()} syntax, which is simpler than the \{\code{map()} when working with list column. For example, calculating the number of raining days can be done by:

@@

The nested form can be used for those operations whose output is only indexed by the spatial identifier (\code{key}), while for those outputs cross-identified by the spatial and temporal identifier (\code{key} and \code{index}), the nested form becomes inadequate. \pkg{cubble} also provides a long form, which expands the \code{ts} column and temporarily ``hide'' the spatial variables. The function \code{face_temporal()} is used to switch a nested cubble into a long one and the first row in Figure \ref{fig:face} illustrates this operation where the focus of the cube now changes from the site-variable face to the time-variable face. This code switches the cubble just created into its long form:

@@

Sometimes, analysts may need to apply some variable transformation that involves both the spatial and temporal variables. An example of this is the transformation of temporal variables into the spatial dimension in glyph maps, which will be elaborated in section \ref{st_transformation}. This type of operation can be seen as flattening, or \emph{unfolding}, the cube into a 2D data frame. Here the function \code{unfold()} moves the spatial variables \code{long} and \code{lat} into the long cubble:

@@

Some readers may raise the question why a new data structure rather than using \code{dplyr::nest_by()} on the combined data to directly create a list-column. While \code{nest_by()} and its pair \code{unnest()} can mimic the nested form in cubble, cubble can be seen as an attempt to arrange variables from different observational units into a single object, specifically for spatio-temporal data. While \texttt{nest\_by()} can recreate the nested form, it can't create a clean slate for temporal variables as the long form. For temporal variables, carrying the duplicated spatial variables in the combined form is not convenient to work with. This is because these spatial variables will lose after \code{summarise()} and one will need to subset the spatial variables again from the original data and make a left join, which tends to create unnecessary intermediate products in the workflow.

@@

As a spatial data object, \code{sf} creates a simple feature geometry list-column (\code{sfc}) in the data frame to provide spatial operations on various geometry types (\code{POINT}, \code{LINESTRING}, \code{POLYGON}, \code{MULTIPOLYGON}, etc). An \code{sf} object can be supplied as the spatial table, as part of the raw data, along with a temporal table, to create a cubble. Once formed in a cubble, methods for the \code{sfc} class can be applied in the nested form. An example of this is shown in \ref{covid}, which also handles the case when the sites in the two tables don't match exactly. A spatial data object with \code{s2} vector can also be the input of the spatial table.

@@

NetCDF data is another format that is commonly used for storing spatio-temporal data. It has two main components: \emph{dimension} for defining the spatio-temporal grid (longitude, latitude, and time) and \emph{variable} that populates the defined grid. Attributes are usually associated with dimensions and variables in the NetCDF format data and a \href{http://cfconventions.org/}{metadata convention for climate and forecast} has been designed to standardise the format of the attributes. A few packages in R exists for manipulating NetCDF data and this includes a high-level R interface: \pkg{ncdf4} (\protect\hyperlink{ref-ncdf4}{Pierce 2019}), a low-level interface that calls C-interface: \pkg{RNetCDF} (\protect\hyperlink{ref-rnetcdf}{Michna and Woods 2021}, \protect\hyperlink{ref-michna2013rnetcdf}{2013}), and a tidyverse implementation: \pkg{tidync} (\protect\hyperlink{ref-tidync}{Sumner 2020}).

@@

The memory limit with NetCDF data in cubble depends on the number of longitude grid points x the number of latitude grid points x the number of time grid points x the number of variables. Cubble can handle slightly more than 300 x 300 (longitude x longitude) grid points for daily 3 variables in one year. A 300 by 300 spatial grid can be a bounding box of {[}100, -80, 180, 0{]} at 0.25 degree resolution or a global bounding box {[}-180, -90, 180, -90{]} at 1 degree resolution. The spatial grid can also be subsetted to trade for longer time periods and more variables through \code{long_range} and \code{lat_range} if the NetCDF file has finer resolution than needed:

@@

Spatial locations can have grouping structure either inherent in the data or formed by clustering. Rather than analysing variables in the site level, summarised variables in the cluster level can give a crisper picture of local areas. In cubble, \code{switch_key()} can be used to create a new level of grouping of spatial locations by specifying a clustering variable. Figure \ref{fig:illu-hier} illustrates the relationship of cubbles at station and cluster level, in both the long and nested form. By specifying \code{cluster_nested <- station_nested \%>\% switch_key(key = cluster)}, the cubble re-defines the cubble key from \code{id} in \code{station_nested} to \code{cluster} in \code{cluster_nested}. All the spatial variables variant to \code{cluster} are now nested into a \code{.val} column and cluster level variables can be computed in the same fashion as station level variables in \code{station_nested}.
